Windowed residual-minimization model reduction for dynamical systems

### 2019-10-02 main.tex

%\title{Model Reduction via \methodName}
%\title{Windowed least--squares reduced-order models for dynamical systems}
\title{Windowed Residual-Minimization Model Reduction for Dynamical Systems}

@@

The LSPG approach sequentially performs residual minimization over a single
time-step, and as such is not associated with residual minimization over the
entire time interval. \KTC{This is what ST-LSPG does. Why do we ignore it? the
full dynamical system} Indeed, LSPG is not equipped with \textit{a priori}
guarantees of stability or convergence, even for linear time-invariant
systems~\cite{bui_thesis}. Further, LSPG performs residual minimization over a
time step, its performance is affected by the time-step size in two distinct
ways: changing the time-step size modifies the window over which the residual
is minimized \textit{as well as} the time-discretization error. As a result,
LSPG often yields the lowest error for an intermediate value of the time step.
For too small a time-step, LSPG yields results similar to the Galerkin
approach; in the limiting case of $\Delta t \rightarrow 0$, LSPG recovers
Galerkin. For too-large a time-step, the accuracy of LSPG degrades. It is
unclear if this is due to the numerical error associated with enlarging the
time-step size, or rather if it is due to the size of the window the residual
is being minimized over. The LSPG approach is additionally sensitive to the
time discretization scheme. When used with explicit temporal discretization,
for instance, LSPG again reverts to Galerkin. To gain the enhanced stability
associated with LSPG, the underlying time-discretization scheme must be
implicit with an appropriate time-step. This can increase the cost of the
reduced-order model, especially in the case where the optimal time-step
required by LSPG is small.

@@

This work seeks to overcome the practical challenges associated with existing
residual minimization approaches for reduced-order models. To this end, we
develop a reduced-order modeling approach that minimizes the FOM ODE residual
over a series of time windows. The key novelties of the work are two fold.
First, the residual minimization problem is formulated to minimize the
\textit{time-continuous} residual. By working from a continuous viewpoint, the
present formulation decouples the underlying temporal discretization scheme
from the residual minimization problem; thus addressing a key deficiency of
LSPG. The continuous viewpoint additionally exposes two differing methods that
can be used to solve the minimization problems, namely \textit{discretize and
then optimize} (direct) methods and \textit{optimize and then discretize}
(indirect) methods. The second novelty is that the residual is minimized over
\textit{time windows}, as opposed to over a single time step or the entire
time domain. This allows for maximum flexibility in terms of computational
cost and numerical accuracy. We demonstrate that this formulation recovers the
Galerkin, LSPG, and space--time LSPG methods. Figure~\ref{fig:flowchart}
depicts how the proposed method interfaces with existing approaches.

@@

\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.1cm 0cm 1cm 0cm},clip,width=0.99\textwidth]{diagram.pdf} 
\caption{Flow chart of \methodAcronym.} 
\label{fig:flowchart} 
\end{centering} 
\end{figure}

>>>>

%\title{Model Reduction via \methodName}
%\title{Windowed least--squares reduced-order models for dynamical systems}
\title{Windowed residual-minimization model reduction for dynamical systems}

@@

Residual minimization methods were extended-to and developed-for reduced-order
models by a variety of researchers. Similar to LSFEM, residual minimization
approaches for reduced-order models comprise finding the solution within a
low-dimensional trial subspace that minimizes the residual norm of the
full-order model. Residual minimization approaches were first applied to
static (i.e., no time-dependence) reduced-order models
in~\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
and then extended to dynamical systems
in~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
Specifically,
Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
formulated the residual minimization problem for dynamical systems by
sequentially minimizing the \textit{fully discrete} residual (i.e., the
residual after temporal discretization) at each time-step instance. This
formulation is most commonly known as the \textit{Least--Squares
Petrov--Galerkin} (LSPG) method. Ref.~\cite{carlberg_lspg_v_galerkin} provided
numerous analyses on this formulation and examined connections with the
standard Galerkin method and Petrov--Galerkin methods. Critically, it was
shown that (1.) for various time-stepping methods LSPG has an equivalent
continuous form that corresponds to a Petrov--Galerkin projection and (2.) the
standard Galerkin method minimizes the time-instantaneous residual of the
full-order model. Numerous numerical experiments have demonstrated that LSPG
yields more stable (although not always more accurate) solutions than the
Galerkin
method~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
The common physical explanation for this improved performance is that, by
minimizing the fully discrete residual, LSPG minimizes the residual over a
\textit{finite-sized time-window}, rather than instantaneously as is done in
the Galerkin method.

\KTC{Let's discuss. What specifically do we want to point out about
deficiencies in LSPG in this section? Should we also look in ST-LSPG? It seems you're trying
to say that we don't like how the time step size \textit{and} time
discretization method dictate so much about the method (I agree with this). If
so, we should structure this to include other discretize-then-optimize methods
I think. Or we could just focus on the time-step aspect (i.e., not the time
discretization method), in which case we should focus on LSPG.}
\KTC{HERE}

The LSPG approach sequentially performs residual minimization over a single
time-step, and as such is not associated with residual minimization over the
entire time interval. \KTC{This is what ST-LSPG does. Why do we ignore it? the
full dynamical system} Indeed, LSPG is not equipped with \textit{a priori}
guarantees of stability or convergence, even for linear time-invariant
systems~\cite{bui_thesis}. Further, the performance of LSPG depends on the
time step in two distinct ways: changing the time-step size modifies both the
time window over which LSPG minimizes the residual \textit{as well as} the
time-discretization error of the full-order model on which LSPG is based. As a
result, LSPG often yields the smallest error for an intermediate value of the
time step (see, e.g., Ref.~\cite[Figure 9]{carlberg_lspg_v_galerkin}). In
particular, as the time step decreases, the accuracy of LSPG approaches that
of Galerkin (which is often poor)\footnote{This occurs because LSPG projection
is equivalent to Galerkin projection in the limit $\Delta t \rightarrow 0$.};
as the time step increases, the time discretization error of the full-order
model (on which the LSPG ROM is based) increases.  LSPG projection additionally is
additionally sensitive to the time discretization scheme. When used with
explicit temporal discretization, for instance, LSPG again reverts to
Galerkin. To gain the enhanced stability associated with LSPG, the underlying
time-discretization scheme must be implicit with an appropriate time-step.
This can increase the cost of the reduced-order model, especially in the case
where the optimal time-step required by LSPG is small.

@@

This work seeks to overcome the practical challenges associated with existing
residual minimization approaches for reduced-order models. To this end, we
propose the \methodName\ method for model reduction, which minimizes the FOM ODE residual
over a sequence of time windows. The key novelties of the work are two fold.
First, the residual minimization problem is formulated to minimize the
\textit{time-continuous} residual. By working from a time-continuous viewpoint, the
present formulation decouples the underlying temporal discretization scheme
from the residual minimization problem; thus addressing a key deficiency of
LSPG \KTC{This is also true for ST-LSPG and maybe other discretization-based
methods? Have to check}. The time-continuous viewpoint additionally exposes
two different methods that
can be used to solve the residual-minimization problems, namely
a \textit{discretize-then-optimize} (i.e., direct) method and
an \textit{optimize-then-discretize} (i.e., indirect) method. The second novelty
is that the residual is minimized over \textit{time windows} rather than over
a single time step or the entire time domain. This equips the method with
additionaly flexibility that enables it
to explore more fine-grained tradeoffs in computational cost and error. We
demonstrate that this method can be viewed as a generalization of existing
methods, as particular instances of this formulation recover Galerkin, LSPG,
and space--time LSPG projection. Figure~\ref{fig:flowchart} depicts how the
proposed method is related to existing approaches.

@@

\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.1cm 0cm 1cm 0cm},clip,width=0.99\textwidth]{diagram.pdf} 
\caption{Relationship diagram for the \methodAcronym\ framework for model
	reduction.} 
\label{fig:flowchart} 
\end{centering} 
\end{figure}
